Study of the numerical solution of X-Modes equations around the hybrid resonance

### 2015-01-29 report/numerics.tex
detail : changed align* command once
changed the pic to a cooler one, removed incorrect data (due to a bug in the code)

We use the following well-known facts:
\begin{itemize}
 \item The C\'ea's lemma applied to the problem (\ref{vf1dcase}); here $C_c$ is the continuity and $C_i$ is the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{h,\nu}\|_{V}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}
 \leq \nu^{-1}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness}.  
\item The form of the exact solution to the problem (\ref{vf1dcase}):
\begin{align}
\label{eq:exact}
 E_x^{\nu}=-iE_{y}^{\nu}\frac{\delta}{\alpha+i\nu}=\frac{f(x)}{\alpha(x)+i\nu},
\end{align}
for some $f(x)\in L_{2}(\Omega)$. 
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:bsc}
 \|v-I^{h}v\|_{L_{2}(\Omega)}\leq Ch^2\left|v''\right|_{L_{2}(\Omega)},\; C>0,
 %\\ \|v-I^{h}v\|_{H_{1}(\Omega)}\leq Ch|v''|_{L_{2}(\Omega)},\; C>0, 
\end{align}
where $I^{h}v$ is an interpolation operator onto $P_{h}^{1}$.
\end{itemize}
For $E_x^{\nu},\;f(x)$ in (\ref{eq:exact}) being sufficiently smooth, 
\ben
 \frac{d^2}{dx^2}E_x^{\nu}=\frac{f''}{\alpha+i\nu}-2\frac{f'\alpha'}{(\alpha+i\nu)^2}+\frac{f\alpha''}{(\alpha+i\nu)^3},
\een
from which, together with (\ref{eq:cond}), it follows that there exists $c>0$ s.t. for all sufficiently small $\nu$ 
\begin{align*}
 \left|\frac{d^2}{dx^2}E_x^{\nu}\right|_{L_2}^{2}&\leq c\int\limits_{\Omega}\frac{1}{(x^2+\nu^2)^{3}}dx
  &=c\left.
 \left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\end{align*}
where $C>0$ does not depend on $\nu$. This, together with the estimates (\ref{eq:cea}) and (\ref{eq:bsc}), results in 
\ben
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}\leq C\nu^{-\frac{7}{2}}h^2,
\een
from which it follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$. Let us check optimality of this estimate.  
To do so, we conduct the  numerical experiment with parameters for the problem as given in Table \ref{tab:parameters}.
\begin{table}
\begin{tabular}{l|r}
Parameter & Value \\
\hline
$\alpha(x)$ & $\left\{\begin{array}{cc}
10, & x\leq -10,\\
-x, & -10<x\leq 5,\\
-5, & x>5.
\end{array}\right.$ \\
$\delta(x)$ & 
$\left\{\begin{array}{cc}
0, & x\leq -10,\\
4/30x+4/3,& -10<x\leq 5,\\
2, & x>5.
\end{array}\right.$ \\
$g^{inc}(-L)$ & $-2 \sqrt{2}i\exp(-22\sqrt{2}i)$\\
$\lambda$ & 
$\sqrt{10}$\\
$L$& 15\\
$H$ & 10 \\
\end{tabular}
\caption{The parameters for the problem with the resonance}
\label{tab:parameters}
\end{table}
We fix $\nu$ and compute the error of the solution compared to the solution computed on the finest mesh  (with $h=2e-6$). 
 The results are depicted in Fig. \ref{fig:dependence} and \ref{fig:l2_err_e2}.
\begin{figure}
\begin{tabular}{cc}
\begin{tikzpicture}
    \begin{loglogaxis}[
    	width=0.45\textwidth,
    	height=0.5\textwidth,
        xlabel=$h$,
        ylabel=$Error$,
        ymin=1e-10,
        legend style={
at={(0.2,1.5)},
legend pos= south east,
anchor=south east, 
draw=none, 
font=\tiny
}
]
\addplot[mark=*,blue] table {pics/result_l2e1.1.dat}; 
\addlegendentry{$\nu=1$};
\addplot[mark=*,green] table {pics/result_l2e1.0.25.dat}; 
\addlegendentry{$\nu=2^{-2}$};
\addplot[mark=*,red] table {pics/result_l2e1.0.0625.dat}; 
\addlegendentry{$\nu=2^{-4}$};
\addplot[mark=square,blue] table {pics/result_l2e1.0.015625.dat}; 
\addlegendentry{$\nu=2^{-6}$};
\addplot[mark=square,green] table {pics/result_l2e1.0.0039062.dat}; 
\addlegendentry{$\nu=2^{-8}$};
\addplot[mark=triangle,red] table {pics/result_l2e1.0.00097656.dat}; 
\addlegendentry{$\nu=2^{-10}$};
\addplot[mark=triangle,blue] table {pics/result_l2e1.0.00024414.dat}; 
\addlegendentry{$\nu=2^{-12}$};
\addplot[mark=triangle, green] table {pics/result_l2e1.6.1035e-05.dat};
\addlegendentry{$\nu=2^{-14}$};
\addplot[mark=triangle, red] table{pics/result_l2e1.1.5259e-05.dat};
\addlegendentry{$\nu=2^{-16}$};
\end{loglogaxis}
\end{tikzpicture} &
\begin{tikzpicture}
    \begin{loglogaxis}[
    	width=0.45\textwidth,
    	height=0.5\textwidth,
        xlabel=$h$,
        ylabel=$Error$,
        ymin=1e-8,
        legend style={
at={(0.2,1.5)},
legend pos= south east,
anchor=south east, 
draw=none,
font=\tiny
}
]
\addplot[mark=*,blue] table {pics/result_l2e2.1.dat}; 
\addlegendentry{$\nu=1$};
\addplot[mark=*,green] table {pics/result_l2e2.0.25.dat}; 
\addlegendentry{$\nu=2^{-2}$};
\addplot[mark=*,red] table {pics/result_l2e2.0.0625.dat}; 
\addlegendentry{$\nu=2^{-4}$};
\addplot[mark=square,blue] table {pics/result_l2e2.0.015625.dat}; 
\addlegendentry{$\nu=2^{-6}$};
\addplot[mark=square,green] table {pics/result_l2e2.0.0039062.dat}; 
\addlegendentry{$\nu=2^{-8}$};
\addplot[mark=square, red] table {pics/result_l2e2.0.00097656.dat}; 
\addlegendentry{$\nu=2^{-10}$};
\addplot[mark=triangle,blue] table {pics/result_l2e2.0.00024414.dat}; 
\addlegendentry{$\nu=2^{-12}$};
\addplot[mark=triangle, green] table{pics/result_l2e2.6.1035e-05.dat};
\addlegendentry{$\nu=2^{-14}$};
\addplot[mark=triangle, red] table{pics/result_l2e2.1.5259e-05.dat};
\addlegendentry{$\nu=2^{-16}$};
\end{loglogaxis}
\end{tikzpicture} 
\end{tabular}
\caption{In the left figure we show the $L_{2}$-error of $E_{x}(x)$, and in the right figure the $L_2$-error of $E_y(x)$ is shown.}
\label{fig:l2_err_e2}
\end{figure}

Denoting by $E_{x}^{h}$ the solution $E_{x}(x)$ computed on the mesh with a width $h$, and by $E_{x}^{c}$ the solution computed on the finest mesh, let 
\begin{align}
\label{eq:def_epsilon}
h_{\epsilon}=\sup\{h: \|E_{x}^{h'}-E_{x}^{c}\|<\epsilon \text{ for all } h'<h\}.
\end{align}
The computed dependence of $h_{\epsilon}$ on $\nu$ is shown in Fig.~\ref{fig:dependence}.
\begin{figure}
\begin{tabular}{cc}
\begin{tikzpicture}
    \begin{loglogaxis}[
    	width=0.4\textwidth,
    	height=0.4\textwidth,
        xlabel=$\nu$,
        ylabel=$h_{\epsilon}$,
        legend style={
at={(0.2,1.5)},
legend pos=north west,
%anchor=south east, 
draw=none
}
]
\addplot[mark=*,blue] table {pics/h1e-1.dat}; 
\addlegendentry{$\epsilon=1e-1$};
%\addplot[mark=o,red] table {pics/h1e-2.dat}; 
%\addlegendentry{$\epsilon=1e-2$};
\addplot[mark=triangle,green] table {pics/h1e-3.dat}; 
\addlegendentry{$\epsilon=1e-3$};
%\addplot[mark=square,cyan] table {pics/h1e-4.dat}; 
%\addlegendentry{$\epsilon=1e-4$};
\addplot[dashed, red] table {pics/nu1.dat}; 
\addlegendentry{$O(\nu)$};
\addplot[dashed, blue] table {pics/nu32.dat}; 
\addlegendentry{$O(\nu^{\frac{3}{2}})$};
\end{loglogaxis}
\end{tikzpicture} 
&
\includegraphics[width=0.4\textwidth]{figure1-crop.pdf}
\end{tabular}
\caption{In the left figure the depenence of $h_{\epsilon}$ as defined by (\ref{eq:def_epsilon}) on $\nu$ is demonstrated. 
In the right figure we show the computed solutions for the problem with parameters in Table \ref{tab:parameters} for different values of $\nu$. }
\label{fig:dependence}
\end{figure}
We can see that the estimate (\ref{eq:estimate_h}) is pessimistic compared to the one suggested by (\ref{eq:def_epsilon}), 
at least for a given value of $\epsilon$.

>>>>

We use the following well-known facts:
\begin{itemize}
 \item The C\'ea's lemma applied to the problem (\ref{vf1dcase}); here $C_c$ is the continuity and $C_i$ is the coercivity constants:
\begin{align}
\label{eq:cea}
\begin{split}
 \|\mathbf{E}^{\nu}-\mathbf{E}^{h,\nu}\|_{V}\leq \frac{C_c}{C_i}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}
 \leq \nu^{-1}\min_{\mathbf{v}\in V_h}\|\mathbf{E}-\mathbf{v}\|_{V}.
 \end{split}
\end{align}
The last inequality follows from Lemma \ref{lemma:well_posedness}.  
\item The form of the exact solution to the problem (\ref{vf1dcase}):
\begin{align}
\label{eq:exact}
 E_x^{\nu}=-iE_{y}^{\nu}\frac{\delta}{\alpha+i\nu}=\frac{f(x)}{\alpha(x)+i\nu},
\end{align}
for some $f(x)\in L_{2}(\Omega)$. 
\item The estimate from \cite[Chapter 0]{brenner} on the rate of convergence of the interpolation 
\begin{align}
\label{eq:bsc}
 \|v-I^{h}v\|_{L_{2}(\Omega)}\leq Ch^2\left|v''\right|_{L_{2}(\Omega)},\; C>0,
 %\\ \|v-I^{h}v\|_{H_{1}(\Omega)}\leq Ch|v''|_{L_{2}(\Omega)},\; C>0, 
\end{align}
where $I^{h}v$ is an interpolation operator onto $P_{h}^{1}$.
\end{itemize}
For $E_x^{\nu},\;f(x)$ in (\ref{eq:exact}) being sufficiently smooth, 
\ben
 \frac{d^2}{dx^2}E_x^{\nu}=\frac{f''}{\alpha+i\nu}-2\frac{f'\alpha'}{(\alpha+i\nu)^2}+\frac{f\alpha''}{(\alpha+i\nu)^3},
\een
from which, together with (\ref{eq:cond}), it follows that there exists $c>0$ s.t. for all sufficiently small $\nu$ 
\bealn
 \left|\frac{d^2}{dx^2}E_x^{\nu}\right|_{L_2}^{2}&\leq c\int\limits_{\Omega}\frac{1}{(x^2+\nu^2)^{3}}dx
  &=c\left.
 \left(\frac{x}{4(x^2+\nu^2)^2}\nu^{-2}+\frac{3x}{8(x^2+\nu^2)}\nu^{-4}+\frac{3}{8}\nu^{-5}\operatorname{atan}\frac{x}{\nu}\right)\right|_{-L}^{H}
 \leq C\nu^{-5},\; 
\eealn
where $C>0$ does not depend on $\nu$. This, together with the estimates (\ref{eq:cea}) and (\ref{eq:bsc}), results in 
\ben
 \|E^{\nu}_{x}-E^{\nu,h}_{x}\|_{L_{2}(\Omega)}\leq C\nu^{-\frac{7}{2}}h^2,
\een
from which it follows that to ensure (\ref{eq:problem1}) $h$ should be chosen as 
\begin{align}
\label{eq:estimate_h}
 h=\alpha_{\epsilon}\nu^{\frac{7}{4}},
\end{align}
where $\alpha_{\epsilon}>0$ depends on $\epsilon$ but does not depend on $\nu$. Let us check optimality of this estimate.  
To do so, we conduct the  numerical experiment with parameters for the problem as given in Table \ref{tab:parameters}.
\begin{table}
\begin{tabular}{l|r}
Parameter & Value \\
\hline
$\alpha(x)$ & $\left\{\begin{array}{cc}
10, & x\leq -10,\\
-x, & -10<x\leq 5,\\
-5, & x>5.
\end{array}\right.$ \\
$\delta(x)$ & 
$\left\{\begin{array}{cc}
0, & x\leq -10,\\
4/30x+4/3,& -10<x\leq 5,\\
2, & x>5.
\end{array}\right.$ \\
$g^{inc}(-L)$ & $-2 \sqrt{2}i\exp(-22\sqrt{2}i)$\\
$\lambda$ & 
$\sqrt{10}$\\
$L$& 15\\
$H$ & 10 \\
\end{tabular}
\caption{The parameters for the problem with the resonance}
\label{tab:parameters}
\end{table}
The results of the experiments for different $\nu$ are shown in Figure \ref{fig:dependence}.

Denoting by $E_{x}^{h}$ the solution $E_{x}(x)$ computed on the mesh with a width $h$, and by $E_{x}^{c}$ the solution computed on a \urev{mesh multiple times finer}, let 
\begin{align}
\label{eq:def_epsilon}
h_{\epsilon}=\sup\{h: \|E_{x}^{h'}-E_{x}^{c}\|<\epsilon \text{ for all } h'<h\}.
\end{align}
The computed dependence of $h_{\epsilon}$ on $\nu$ is shown in Fig.~\ref{fig:dependence}.
\begin{figure}
\begin{tabular}{cc}
\begin{tikzpicture}
    \begin{loglogaxis}[
    	width=0.4\textwidth,
    	height=0.4\textwidth,
        xlabel=$\nu$,
        ylabel=$h_{\epsilon}$,
        legend style={
at={(0.2,1.5)},
legend pos=north west,
%anchor=south east, 
draw=none, 
xmin=0.5e-5
}
]
\addplot[mark=*,blue] table {pics/h1e-1.dat}; 
\addlegendentry{$\epsilon=1e-1$};
%\addplot[mark=o,red] table {pics/h1e-2.dat}; 
%\addlegendentry{$\epsilon=1e-2$};
\addplot[mark=triangle,green] table {pics/h1e-3.dat}; 
\addlegendentry{$\epsilon=1e-3$};
%\addplot[mark=square,cyan] table {pics/h1e-4.dat}; 
%\addlegendentry{$\epsilon=1e-4$};
\addplot[dashed, red] table {pics/nu1.dat}; 
\addlegendentry{$O(\nu)$};
\addplot[dashed, blue] table {pics/nu32.dat}; 
\addlegendentry{$O(\nu^{\frac{3}{2}})$};
\end{loglogaxis}
\end{tikzpicture} 
&
\includegraphics[width=0.4\textwidth]{figure1-crop.pdf}
\end{tabular}
\caption{In the left figure the depenence of $h_{\epsilon}$ as defined by (\ref{eq:def_epsilon}) on $\nu$ is demonstrated. 
In the right figure we show the computed solutions for the problem with parameters in Table \ref{tab:parameters} for different values of $\nu$. }
\label{fig:dependence}
\end{figure}
We can see that the estimate (\ref{eq:estimate_h}) is pessimistic compared to the one suggested by (\ref{eq:def_epsilon}), 
at least for a given value of $\epsilon$.
